Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-10-26 writedown/period.noneq.msm/paper.tex
add intro

\section{Introduction}
Non-equilibrium, especially periodically driven system. Interesting.

MSM tools for analyzing, provide profound understanding.

Current achievement of MSM in equilibrium cases:
\begin{itemize}
\item well developed in theory and applications \cite{A19-31, A19-1}
\item software for MSM building \cite{A19-49, MSMBuilder}
\end{itemize}

Importance: first application of MSM in a non-equilibrium system.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
For the sake of simplicity we consider diffusive molecular dynamics in an energy landscape $V$ driven by the time-dependet external driving force $E(t)D(x_t)$ with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t)
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p(i,t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ componentwise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

Since $P$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with right eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a left eigenvector $\mymu$
satisfying
$\mymu^{\top}P=\mymu^{\top}$.
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_i\mu(i)=1$). In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

@@

MSM building has been attracted a lot of attention recently, and theory  \cite{A19-31} as well as algorithms \cite{A19-1}, applications (see e.g. \cite{A19-26,PNAS09} for two of hundreds of articles) and software \cite{A19-49, MSMBuilder} have been developed to quite an extend. However, 
by far most of the literature is related to building \emph{standard} MSMs for equilibirium MD. In standard MSM also the transition region has to be discretized, a feature that often forces the user to incorporate more macrostates than essentially needed to approximate the long-term kinetics.
In Ref.~\cite{sarich2010approximation, A19-31,schuette2011markov,BucheteHummer} it has been shown how to construct \emph{non-standard} MSM that avoid this problem for equilibrium MD, i.e., if $\vect P$ satisfies the detailed balance
condition: (1) Identify the cores of the metastable sets of the
dynamics, (2) use them as milestones to construct an MSM in which the
macrostates are the metastable core sets and $\hat{\vect P}$ is the transition
matrix of the milestone process \cite{A19-31,schuette2011markov,A19-29} that models the jumping behavior of
the original dynamics between the metastable regions.

@@

We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ can be computed by matrix multiplication using the committors:
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}_j}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$ (for appropriately chosen core sets).

@@

If all of its entries are positive such that it is a stochastic matrix, $\hat{M}^{-1}\hat{T}$ thus can also be taken as MSM transition matrices. In the case of $\hat{M}^{-1}\hat{T}$ the MSM modelling error results from Galerkin discretization, while the MSM modelling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibirium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactoins,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of length $10^6$~ps, along which snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching NEMD
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

For this periodically driven molecular system we will first show how to choose an appropriately fine discretization. After validating this discretization we will consider
the discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally we will coarse grain this description further by constructiion of a $3$ state Markov State Model that is able to describe the long-term kinetics of the system correctly.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^{\top}(T)$ just by counting the transition behavior of MD trajectories 
\begin{equation}\label{Pdirect}
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\end{equation}
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuos description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time.
  Since the dimensionality is reduced by using only a few collective
  variables, the lag-time should be chosen large enough so that
  the original dynamics $x_t$ is properly relaxed with regard to
  the unresolved degrees of freedoms on timescales shorter than the lag-time
  (assuming  that the collective variables capture the slow dynamics).
  In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{fig/t010/cluster.marco.3/fig-coreset-prob.eps}
  \caption{Comparison of NEMD and MSM long-term kinetics of alanine
    dipeptide in an oscillatory electrc field. The plots show the
    time-dependent probability $\hat p_j(t)$ to be assigned to
    core set $C_j$ (corresponding to the observable $\mathcal A$ given
    in (\ref{A}) with $\alpha_j=1$ and $\alpha_k=0$ for $k\not=
    j$). Solid lines are from brute force NEMD simulations, while the
    dashed lines are from our MSM.}
  \label{fig:num-7}
\end{figure}

@@

\section{Concluding remarks and discussions}

>>>>

\section{Introduction}
% \textbf{Non-equilibrium, especially periodically driven system. Interesting.}

The non-equilibrium biomolecular systems, especially
those systems under periodic driven, are attracting increasing
interest.  For example, the potential effects of an electromagnetic
radiation on the human body tissue (e.g.~DNA, protein, and membrane)  has been extensively
investigated by a vast number of literatures, and a very limited
subset of them is Ref.~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects,
  wang2014exploring}.  Among those theoretical investigations, the
molecular dynamics (MD) has been shown  useful in understanding
the response of the biomolecular conformation to the external electric field,
due to its ability of resolving the molecular
details, which are sometimes not achievable in experiments. Only
recently, a dynamical non-equilibrium MD simulation (D-NEMD) method has
been proposed.  With the local thermostating technique, the D-NEMD
method is able to sample the non-equilibrium probability density at the
constant temperature, therefore opens the possibility of studying the
non-equilibrium processes in a statistically reliable and
thermodynamically consistent way.

% \textbf{MSM tools for analyzing, provide profound understanding.}

Despite their high resolution in presenting a molecular system, MD the
trajectories including the time-dependent position and velocity of every atom
sometimes impose non-trivial challenge for data analyzing, in the sense that 
extracting the essential structure and dynamical properties is
becoming increasingly difficult as the  system size and trajectory length grow.
% by merely visulizing (not only molecular cartoons but also order parameters)
% and observing the molecules.
The Markov state model
(MSM)~\cite{A19-31}, which is a promising method for data analyzing of
the high-dimensional MD trajectories, has been well developed over the
past decade in theory and applications \cite{A19-31, A19-1}, and in
the software implementations~\cite{A19-49, MSMBuilder}.  The principal
idea of MSM is to approximate the original high-dimensional system by
a reduced Markovian dynamics over a finite number of states, from
which the essential structures and dynamics can be formulated by the spectrum analysis.
By assigning the spacial and temporal discretizations, 
the transition probabilities are sampled from MD trajectories, and
the MSM does not need more information from the MD simulation.
The quality of the MSM therefore relies on the quality of the discretizations.
Moreover, the characteristic of the original dynamics plays a role:
It has been shown that in principle good approximate accuracy would be achieved
by using only a few ``coresets'',
if the system is dominated by a few metastable states, in the sense that
the MD trajectories stay in the vicinity of a metastable state substantially longer than
transform to another such state~\cite{schuette2011markov}.
Besides the applications in analyzing the MD trajectories, it worth noting that MSM is
also used for designing initial configurations for massively parallel MD
simulations~\cite{kohlhoff2014cloud}.

% Current achievement of MSM in equilibrium cases:
% \begin{itemize}
% \item well developed in theory and applications \cite{A19-31, A19-1}
% \item software for MSM building \cite{A19-49, MSMBuilder}
% \end{itemize}

% \textbf{Importance: first application of MSM in a non-equilibrium system.}

Despite the significance of the non-equilibrium phenomena, the
analysis of the non-equilibrium MD simulations mainly follows the
visualizing (not only molecular cartoons but also order parameters)
and observing approach, and the quantitative description of
the essential conformational dynamics is still lacking. On the other
hand, the MSM, to the best of our knowledge, has only been applied to
equilibrium systems,  and most of them are assumed reversible. This work is the first attempt
of using the MSM in analyzing a class of truly non-equilibrium systems:
periodic external driven. Being more specifically,
we investigate the protein conformational dynamics
under the oscillating electric field (EF).
In Sec.~\ref{sec:disc}, we discuss the temporal and spacial
discretizations of the non-equilibrium dynamics. The spacial discretization
is on the dihedral angle space, which can be interpreted as a
MSM in the traditional sense of full spacial partition.
In the temporal direction, the non-equilibrium process
is  discretized by the Floquet theorem into a time-homogeneous
but not necessarily reversible dynamics. Since the full spacial partition
is not preferable in high-dimensional system, in Sec.~\ref{sec:build-msm},
the MSM based on milestoning is introduced for the discretized irreversible process.
The validity of the discretizations and the MSM is checked in Sec.~\ref{sec:alanine}
by comparing to the brute force
non-equilibrium simulation of alanine dipeptide under oscillating EF.
This work is summarized in Sec.~\ref{sec:conclusion} by a chart presenting
the workflow of building MSM for non-equilibrium systems,
and open questions are presented in the end.

@@

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
For the sake of simplicity we consider diffusive molecular dynamics in an energy landscape $V$ driven by the time-dependent external driving force $E(t)D(x_t)$ with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibrium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t)
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p(i,t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p(i,0)\ge 0$ componentwise, then $p(i,t)\ge 0$ and $\sum_i
p(i,t) = \sum_ip(i,0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

Since $P$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially complex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with right eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a left eigenvector $\mymu$
satisfying
$\mymu^{\top}P=\mymu^{\top}$.
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_i\mu(i)=1$). In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

@@

MSM building has been attracted a lot of attention recently, and theory  \cite{A19-31} as well as algorithms \cite{A19-1}, applications (see e.g. \cite{A19-26,PNAS09} for two of hundreds of articles) and software \cite{A19-49, MSMBuilder} have been developed to quite an extend. However, 
by far most of the literature is related to building \emph{standard} MSMs for equilibrium MD. In standard MSM also the transition region has to be discretized, a feature that often forces the user to incorporate more macrostates than essentially needed to approximate the long-term kinetics.
In Ref.~\cite{sarich2010approximation, A19-31,schuette2011markov,BucheteHummer} it has been shown how to construct \emph{non-standard} MSM that avoid this problem for equilibrium MD, i.e., if $\vect P$ satisfies the detailed balance
condition: (1) Identify the cores of the metastable sets of the
dynamics, (2) use them as milestones to construct an MSM in which the
macrostates are the metastable core sets and $\hat{\vect P}$ is the transition
matrix of the milestone process \cite{A19-31,schuette2011markov,A19-29} that models the jumping behavior of
the original dynamics between the metastable regions.

@@

We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ can be computed by matrix multiplication using the committors:
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}_j}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timescale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modeling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$ (for appropriately chosen core sets).

@@

If all of its entries are positive such that it is a stochastic matrix, $\hat{M}^{-1}\hat{T}$ thus can also be taken as MSM transition matrices. In the case of $\hat{M}^{-1}\hat{T}$ the MSM modeling error results from Galerkin discretization, while the MSM modeling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibrium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactions,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of length $10^6$~ps, along which snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching NEMD
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

For this periodically driven molecular system we will first show how to choose an appropriately fine discretization. After validating this discretization we will consider
the discretized dynamics generated by the Floquet transition matrix $\vect P$
in comparison to the original NEMD simulation. Finally we will coarse grain this description further by construction of a $3$ state Markov State Model that is able to describe the long-term kinetics of the system correctly.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^{\top}(T)$ just by counting the transition behavior of MD trajectories 
\begin{equation}\label{Pdirect}
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\end{equation}
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuous description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time.
  Since the dimensionality is reduced by using only a few collective
  variables, the lag-time should be chosen large enough so that
  the original dynamics $x_t$ is properly relaxed with regard to
  the unresolved degrees of freedoms on timescales shorter than the lag-time
  (assuming  that the collective variables capture the slow dynamics).
  In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{fig/t010/cluster.marco.3/fig-coreset-prob.eps}
  \caption{Comparison of NEMD and MSM long-term kinetics of alanine
    dipeptide in an oscillatory electric field. The plots show the
    time-dependent probability $\hat p_j(t)$ to be assigned to
    core set $C_j$ (corresponding to the observable $\mathcal A$ given
    in (\ref{A}) with $\alpha_j=1$ and $\alpha_k=0$ for $k\not=
    j$). Solid lines are from brute force NEMD simulations, while the
    dashed lines are from our MSM.}
  \label{fig:num-7}
\end{figure}

@@

\section{Concluding remarks and discussions}
\label{sec:conclusion}
